Toxic Code Snippets on Stack Overflow

### 2017-11-13 cloverflow-tse/cloverflow.tex
Formatting, small fixes, and modifying the abstract to fit with 200-word TSE limit.

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEtitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\IEEEtitleabstractindextext{%
\begin{abstract}
Online code clones are code fragments that are copied from software projects or
online sources to Stack Overflow as examples in questions and answers. % They
Due to an absence of a checking mechanism after the code has been copied to
Stack Overflow, they can become toxic code snippets, i.e.~they suffer from
being outdated or violating the original software license. We present a study
incorporating several state-of-the-art tools and techniques to automatically
extract and filter detected online clone pairs between 72,365 Java code snippets
on Stack Overflow and 111 Java open source projects in the curated Qualitas
corpus. We discovered and analysed 2,302 non-trivial online clone candidates.
Our investigation of the online clone pairs revealed strong evidence that 154
clones have been copied from a Qualitas project to Stack Overflow.  We found 101
of them (66\%) to be outdated and potentially harmful for reuse.  Furthermore,
we found 214 code snippets on Stack Overflow that could potentially violate the license of their
original software. To better understand the phenomenon, we also surveyed 201 high-reputation Stack Overflow answerers
(33\% response rate) and found that 131 of the answerers (65\%) have ever been notified of
outdated code in the answers and 26 of them (20\%) rarely or never fix the outdated
code. Only 3 answerers explicitly include software licenses in their answers and
138 answerers (69\%) never check for licensing conflicts between their copied code
snippets and Stack Overflow's CC BY-SA 3.0.
\end{abstract}

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
Clone Detection, Stack Overflow, Outdated Code, Software Licensing
\end{IEEEkeywords}}

@@

%The popularity of the Internet encourages tremendous amount of source code being shared online. 
Stack Overflow is a popular online programming community with 7.6
million users, 14 million questions, and 23 million answers\footnote{Data as of 21
	August 2017 from~\url{https://stackexchange.com/sites}}. It allows programmers to ask questions and give answers
to programming problems. The website has found to be useful for
software
development~\cite{Ponzanelli2013,Ponzanelli2014,Keivanloo2014,Park2014,
	Stolee2014,Subramanian2013,Diamantopoulos2015,Treude2016} and also
valuable for educational purposes~\cite{Nasehi2012}. On Stack
Overflow, each conversation contains a question and a list of
answers. The answers frequently contain at least one code snippet as
a solution to the question asked. We found that the code snippets are
usually not authored directly on the Stack Overflow website but copied
from another location. A snippet in an answer could be copied and modified
from a code snippet in the question, copied from the answerer's own
code or from other locations including open source software (OSS)
systems.

@@

\begin{figure*}
	\begin{lstlisting}
/* Code in Stack Overflow post ID 801987 */            /* StringUtils.java (2013-02-05) */
public static String humanReadableInt(long number) {   public static String humanReadableInt(long number) {
  long absNumber = Math.abs(number);                     return TraditionalBinaryPrefix.long2String(number,"",1);
  double result = number;                              }
  String suffix = "";
  if (absNumber < 1024) {
  } else if (absNumber < 1024 * 1024) {
    result = number / 1024.0;
    suffix = "k";
  } else if (absNumber < 1024 * 1024 * 1024) {
    result = number / (1024.0 * 1024);
    suffix = "m";
  } else {
    result = number / (1024.0 * 1024 * 1024);
    suffix = "g";
  }
  return oneDecimal.format(result) + suffix;
}
	\end{lstlisting}\vspace{-2ex}
	\caption{An example of the two code fragments of
		{\small\texttt{StringUtils.java}}. The one from the
		Stack Overflow post 801987 (left) is outdated when compared to
		its latest version in the \textsf{Hadoop} code base
		(right). The toxic code snippet is outdated code, has race conditions, and does not contain its
		original Apache v.2.0 license.}
	\label{fig:before-after_2}
\end{figure*}

The second motivating example of outdated online code clones with more
disrupting changes than the first one  can be found in an answer to a Stack
Overflow question regarding how to format files sizes in a human readable form.
\Cref{fig:before-after_2} shows---on the left---a code snippet to perform the
task from the {\small{\texttt{StringUtils}}} class in \textsf{Hadoop}. The code snippet on the
right shows another version of the same method, but at this time extracted from
the latest version of \textsf{Hadoop}. We can see that
they are totally different. The {\small{\texttt{humanReadableInt}}} method is
rewritten on February 5, 2013 to solve an issue of a race condition
(issue no.~HADOOP-9252\footnote{https://issues.apache.org/jira/browse/HADOOP-9252}).
Similar to the first example, the clone code snippet on Stack Overflow does not 
include its original Apache v.2.0 license.

@@

\subsection{Stack Overflow Developers' Survey} We support our findings of
toxic code snippets on Stack Overflow by asking Stack
Overflow users to take an online survey. The survey was used for assessing awareness of
the developers on the two issues of outdated code and license-violating code
snippets. We designed the survey using Google Forms by following the guidelines by Pfleeger and
Kitchenham~\cite{Pfleeger2001,Kitchenham2002}. The survey was completely
anonymous and the participants could decide to leave at any time. It contained
11 questions. There were 7 Likert's scale questions, 3 yes/no questions, and one
open-ended question for additional comments. The first two questions were
mandatory while the other 9 questions were shown to the participants based on
their previous answers. The full survey can be found in our research note~\cite{Ragkhitwetsagul_RN2017}~\FIXME{Fix the RN bibtex}.

@@

\textbf{IN: Inheritance/interface.} There were 16 clone pairs, 9 pairs after
consolidation, found to be similar because they implement the same interface or
inheriting from the same class. An example is the two implementations of a
custom data type which implements {\small\texttt{UserType}}. They share similar
{\small\texttt{@Override}} methods of {\small\texttt{deepCopy}},
{\small\texttt{isMutable}}, {\small\texttt{assemble}},
{\small\texttt{disassemble}}, and {\small\texttt{replace}}.

@@

\begin{table}
	\centering
	\caption{Stack Overflow answerers taken the survey}
	\label{tab:survey_target}
	\begin{tabular}{rrrr}
		\toprule
		Reputation & Sent emails & Answers & Rate \\
		\midrule
		963,731--6,999 & 607 & 201 & 33\% \\
		\bottomrule
	\end{tabular}
\end{table}

We received 201 answers (33\% response rate) from 607 emails we sent to the
Stack Overflow answerers. The response rate was high considering other
online surveys in software engineering~\cite{Punter2003}. 
We only present a summary of the survey answers in
this paper and the full analysis is available as a technical report~\cite{Ragkhitwetsagul_RN2017}.

\subsubsection{General Information}
As shown in \Cref{tab:survey_exp}, the majority of the answerers are experienced
developers with more than 10 years of experience (55.4\%) or between 5 to 10
years (28.7\%). They are active users and regularly answer questions. 49
participants (24\%) answer questions on Stack Overflow every day.

@@

Work similar to ours are studies by An et al.~\cite{An2017}
Abdalkareem et al.~\cite{Abdalkareem2017}, and Baltes et al.~\cite{Baltes2017}. 
An et al.~investigated clones
between 399 Android apps and Stack Overflow posts. They found 1,226 code
snippets which were reused from 68 Android apps. They also observed that there
are 1,279 cases of potential license violations. The authors rely on the
timestamp to judge whether the code has been copied from/to Stack Overflow along
with confirmations from six developers. Instead of Android apps, we investigated
clones between Stack Overflow and 111 open source projects. Their results are
similar to our findings that there exist clones from software projects to Stack
Overflow with potential licensing violations. Abdalkareem et
al.~\cite{Abdalkareem2017} detected clones between Stack Overflow posts and
Android apps from the F-Droid repository and used timestamps to
determine the direction
of copying. They found 22 apps containing code cloned from Stack Overflow. They
reported that cloned code is commonly used for enhancing existing code. Their
analysis shows that the cloned code from Stack Overflow have negative effects to
quality of the apps. The median percentage of bug fixing commits after adding
Stack Overflow code (19.09\%) is higher than before adding the code (5.96\%)
with statistical significance.
Baltes et al.~\cite{Baltes2017} discovered that two
thirds of the clones from the 10 most frequently referenced Java code snippets on Stack
Overflow do not contain attributions.
%In our work, we defined seven patterns of online code cloning and 
%performed a large-scale manual check of 3,632 clone pairs. 
%We discovered 112 clone pairs with strong evidences, 
%based on natural text in comments and post contents, that they 
%were copied from Qualitas to Stack Overflow. By comparing the 
%clones to their latest versions in the software, we found that 
%64 code snippets on Stack Overflow are outdated and possibly harmful for reuse.

>>>>

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEtitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\IEEEtitleabstractindextext{%
\begin{abstract}
Online code clones are code fragments that are copied from software projects or
online sources to Stack Overflow as examples. Due to an absence of a checking
mechanism after the code has been copied to Stack Overflow, they can become
toxic code snippets, i.e. they suffer from being outdated or violating the
original software license. We present a study incorporating state-of-the-art
tools and techniques to automatically extract and filter online clone pairs
between 72,365 Java code snippets on Stack Overflow and 111 open source projects
in the curated Qualitas corpus. We analysed 2,302 non-trivial online clone
candidates. Our investigation revealed strong evidence that 154 clones have been
copied from a Qualitas project to Stack Overflow. We found 101 of them (66\%) to
be outdated and potentially harmful for reuse. Furthermore, we found 214 code
snippets that could potentially violate the license of their original software.
The survey of 201 high-reputation Stack Overflow answerers (33\% response rate)
showed that 131 participants (65\%) have ever been notified of outdated code and
26 of them (20\%) rarely or never fix the code. 138 answerers (69\%) never check
for licensing conflicts between their copied code snippets and Stack Overflow's
CC BY-SA 3.0.

%Online code clones are code fragments that are copied from software projects or
%online sources to Stack Overflow as examples in questions and answers. % They
%Due to an absence of a checking mechanism after the code has been copied to
%Stack Overflow, they can become toxic code snippets, i.e.~they suffer from
%being outdated or violating the original software license. We present a study
%incorporating several state-of-the-art tools and techniques to automatically
%extract and filter detected online clone pairs between 72,365 Java code snippets
%on Stack Overflow and 111 Java open source projects in the curated Qualitas
%corpus. We discovered and analysed 2,302 non-trivial online clone candidates.
%Our investigation of the online clone pairs revealed strong evidence that 154
%clones have been copied from a Qualitas project to Stack Overflow.  We found 101
%of them (66\%) to be outdated and potentially harmful for reuse.  Furthermore,
%we found 214 code snippets on Stack Overflow that could potentially violate the license of their
%original software. To better understand the phenomenon, we also surveyed 201 high-reputation Stack Overflow answerers
%(33\% response rate) and found that 131 of the answerers (65\%) have ever been notified of
%outdated code in the answers and 26 of them (20\%) rarely or never fix the outdated
%code. Only 3 answerers explicitly include software licenses in their answers and
%138 answerers (69\%) never check for licensing conflicts between their copied code
%snippets and Stack Overflow's CC BY-SA 3.0.
\end{abstract}

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
Code Clone Detection, Stack Overflow, Outdated Code, Software Licensing
\end{IEEEkeywords}}

@@

%The popularity of the Internet encourages tremendous amount of source code
% being shared online.
Stack Overflow is a popular online programming community with 7.6 million users,
14 million questions, and 23 million answers\footnote{Data as of 21 August 2017
	from~\url{https://stackexchange.com/sites}}. It allows programmers to ask
questions and give answers to programming problems. The website has found to be
useful for software
development~\cite{Ponzanelli2013,Ponzanelli2014,Keivanloo2014,Park2014,
	Stolee2014,Subramanian2013,Diamantopoulos2015,Treude2016} and also valuable for
educational purposes~\cite{Nasehi2012}. On Stack Overflow, each conversation
contains a question and a list of answers. The answers frequently contain at
least one code snippet as a solution to the question asked. We found that the
code snippets are usually not authored directly on the Stack Overflow website
but copied from another location. A snippet in an answer could be copied and
modified from a code snippet in the question, copied from the answerer's own
code or from other locations including open source software (OSS) systems.

@@

\begin{figure*}
	\begin{lstlisting}
/* Code in Stack Overflow post ID 801987 */            /* StringUtils.java (2013-02-05) */
public static String humanReadableInt(long number) {   public static String humanReadableInt(long number) {
  long absNumber = Math.abs(number);                     return TraditionalBinaryPrefix.long2String(number,"",1);
  double result = number;                              }
  String suffix = "";
  if (absNumber < 1024) {
  } else if (absNumber < 1024 * 1024) {
    result = number / 1024.0;
    suffix = "k";
  } else if (absNumber < 1024 * 1024 * 1024) {
    result = number / (1024.0 * 1024);
    suffix = "m";
  } else {
    result = number / (1024.0 * 1024 * 1024);
    suffix = "g";
  }
  return oneDecimal.format(result) + suffix;
}
	\end{lstlisting}\vspace{-2ex}
	\caption{An example of the two code fragments of
		{\small\texttt{StringUtils.java}}. The one from the
		Stack Overflow post 801987 (left) is outdated when compared to
		its latest version in the \textsf{Hadoop} code base
		(right). The toxic code snippet is outdated code and has race conditions.}
	\label{fig:before-after_2}
\end{figure*}

The second motivating example of outdated online code clones with more
disrupting changes than the first one  can be found in an answer to a Stack
Overflow question regarding how to format files sizes in a human readable form.
\Cref{fig:before-after_2} shows---on the left---a code snippet to perform the
task from the {\small{\texttt{StringUtils}}} class in \textsf{Hadoop}. The code snippet on the
right shows another version of the same method, but at this time extracted from
the latest version of \textsf{Hadoop}. We can see that
they are totally different. The {\small{\texttt{humanReadableInt}}} method is
rewritten on February 5, 2013 to solve an issue of a race condition
(issue no.~HADOOP-9252\footnote{https://issues.apache.org/jira/browse/HADOOP-9252}).
%Similar to the first example, the clone code snippet on Stack Overflow does not 
%include its original Apache v.2.0 license.

@@

\subsection{Stack Overflow Developers' Survey} We support our findings of
toxic code snippets on Stack Overflow by asking Stack
Overflow users to take an online survey. The survey was used for assessing awareness of
the developers on the two issues of outdated code and license-violating code
snippets. We designed the survey using Google Forms by following the guidelines by Pfleeger and
Kitchenham~\cite{Pfleeger2001,Kitchenham2002}. The survey was completely
anonymous and the participants could decide to leave at any time. It contained
11 questions. There were 7 Likert's scale questions, 3 yes/no questions, and one
open-ended question for additional comments. The first two questions were
mandatory while the other 9 questions were shown to the participants based on
their previous answers. The full survey can be found in our research note~\cite{Ragkhitwetsagul_RN2017}.

@@

\textbf{IN: Inheritance/interface.} There were 16 clone pairs, 9 pairs after
consolidation, found to be similar because they implement the same interface or
inherit from the same class. An example is the two implementations of a
custom data type which implements {\small\texttt{UserType}}. They share similar
{\small\texttt{@Override}} methods of {\small\texttt{deepCopy}},
{\small\texttt{isMutable}}, {\small\texttt{assemble}},
{\small\texttt{disassemble}}, and {\small\texttt{replace}}.

@@

%\begin{table}
%	\centering
%	\caption{Stack Overflow answerers taken the survey}
%	\label{tab:survey_target}
%	\begin{tabular}{rrrr}
%		\toprule
%		Reputation & Sent emails & Answers & Rate \\
%		\midrule
%		963,731--6,999 & 607 & 201 & 33\% \\
%		\bottomrule
%	\end{tabular}
%\end{table}

We received 201 answers (33\% response rate) from 607 emails we sent to the
Stack Overflow answerers. The response rate was high considering other
online surveys in software engineering~\cite{Punter2003}. 
We only present a summary of the survey answers in
this paper and the full analysis is available as a research note~\cite{Ragkhitwetsagul_RN2017}.

\subsubsection{General Information}
As shown in \Cref{tab:survey_exp}, the majority of the answerers are experienced
developers with more than 10 years of experience (55.2\%) or between 5 to 10
years (28.9\%). They are active users and regularly answer questions. 49
participants (24\%) answer questions on Stack Overflow every day.

@@

Work similar to ours are studies by An et al.~\cite{An2017},
Abdalkareem et al.~\cite{Abdalkareem2017}, and Baltes et al.~\cite{Baltes2017}. 
An et al.~investigated clones
between 399 Android apps and Stack Overflow posts. They found 1,226 code
snippets which were reused from 68 Android apps. They also observed that there
are 1,279 cases of potential license violations. The authors rely on the
timestamp to judge whether the code has been copied from/to Stack Overflow along
with confirmations from six developers. Instead of Android apps, we investigated
clones between Stack Overflow and 111 open source projects. Their results are
similar to our findings that there exist clones from software projects to Stack
Overflow with potential licensing violations. Abdalkareem et
al.~\cite{Abdalkareem2017} detected clones between Stack Overflow posts and
Android apps from the F-Droid repository and used timestamps to
determine the direction
of copying. They found 22 apps containing code cloned from Stack Overflow. They
reported that cloned code is commonly used for enhancing existing code. Their
analysis shows that the cloned code from Stack Overflow have negative effects to
quality of the apps. The median percentage of bug fixing commits after adding
Stack Overflow code (19.09\%) is higher than before adding the code (5.96\%)
with statistical significance.
Baltes et al.~\cite{Baltes2017} discovered that two
thirds of the clones from the 10 most frequently referenced Java code snippets on Stack
Overflow do not contain attributions.
%In our work, we defined seven patterns of online code cloning and 
%performed a large-scale manual check of 3,632 clone pairs. 
%We discovered 112 clone pairs with strong evidences, 
%based on natural text in comments and post contents, that they 
%were copied from Qualitas to Stack Overflow. By comparing the 
%clones to their latest versions in the software, we found that 
%64 code snippets on Stack Overflow are outdated and possibly harmful for reuse.
